\chapter{}

The previous proposition will be used to prove 
that the algebraic closure always exists. 

\begin{theorem}[Artin]
	\index{Artin's theorem}
	Let $K$ be a field. Then $K$ admits an algebraic closure $C/K$. If $C_1/K$
	is an algebraic closure, then the extensions $C/K$ and $C_1/K$ are
	isomorphic. 
\end{theorem}

\begin{proof}
    Let us first prove the uniqueness. The previous proposition implies the existence of 
    an extensions homomorphism $\varphi\colon C\to C_1$. Let $y\in C_1$ and $f=f(y,K)$ be 
    the minimal polynomial of $y$ in $K$. Since $f$ admits a factorization
    \[
        f=\lambda\prod (X-\alpha_i)^{m_i}
    \]
    in $C[X]$, it follows that
    \[
    f=\overline{\varphi}(f)=\prod (X-\varphi(\alpha_i))^{m_i}
    \]
    Since $0=f(y)$, we conclude that $y=\varphi(\alpha_j)$ for some $j$. In particular, $\varphi$ is
    surjective and hence $\varphi$ is bijective. 
    
    We now prove the existence. Let us assume that $K$ admits an extension $E/K$ 
    with $E$ algebraically closed. Let 
    \[
    	F=\overline{K}_E=\{x\in E:x\text{ is algebraic over }K\}. 
    \]
    Then $F/K$ is algebraic. Let $g\in F[X]$ be such that
    $\deg g>0$. Since $E$ is algebraically closed, $g$ admits a root $\alpha$ in $E$. In particular, $\alpha$
    is algebraic over $F$ and hence $\alpha$ is algebraic over $K$. This implies that $\alpha\in F$, thus
    $F$ is algebraically closed. This proves that $F/K$ is an algebraic closure. 
    
    Let us prove that there exists an extension $E_1/K$ such that
    every polynomial $f\in K[X]$ with $\deg f>0$ has a root in $E_1$. Let 
    $\{f_i:i\in I\}$ be the family of monic irreducible polynomials with coefficients in $K$. 
    We may think that $f_i=f_i(X_i)$. 
    Let $R=K[\{X_i:i\in I\}]$ and let $J$ be the ideal of $R$ 
    generated by the $f_i(X_i)$. We claim that $J\ne R$. If not, $1\in J$, so
    \[
    1=\sum_{j=1}^m g_jf_{i_j}(X_j)
    \]
    for some $g_1,\dots,g_m\in R$. There exists an extension $F/K$ such that
    $f_{i_j}$ has a root $\alpha_j$ in $F$ for all $j$. Let 
    \[
    \sigma\colon R\to F,\quad
    \sigma(X_k)=\begin{cases}
        \alpha_j & \text{if $k=i_j$},\\
        0 & \text{if $k\not\in\{i_1,\dots,i_m\}$}.
        \end{cases}
    \]
    Then $1=\sigma(1)=\sum_{j=1}^m\sigma(g_j)f_{i_j}(\alpha_j)=0$, a contradiction. 
    
    Since $J$ is a proper ideal, it is contained in a maximal ideal $M$. Let $L=R/M$ 
    and let $\sigma\colon K\to L$ be the composition $K\hookrightarrow R\to R/M=L$. 
    As we did before,
    $\pi(X_i)$ is a root of $\overline{\sigma}(f_i)$ for all $i$n
    and there exists an extension $E_1/K$ such that
    every $f_i$ has a root in $E_1$. Proceeding in this way, we construct
    a sequence
    \[
    E_1\subseteq E_2\subseteq\cdots
    \]
    of fields such that every polynomial of positive degree and coefficients in $E_k$ 
    admits a root in $E_{k+1}$. Let $E=\cup E_k$. We claim that $E$ is algebraically closed. In fact, 
    let $g\in E[X]$ be such that $\deg g>0$. Then, since $g\in E_r[X]$ for some $r$, it follows
    that $g$ has a root in $E_{r+1}\subseteq E$. 
\end{proof}

\topic{Decomposition fields}

\begin{definition}
\index{Decomposition field}
	Let $K$ be a field and $f\in K[X]$ be such that $\deg f>0$. A \textbf{decomposition field}
	of $f$ over $K$ is field $E$ that contains $K$ and that satisfies the following properties:
	\begin{enumerate}
		\item $f$ factorizes linearly in $E[X]$. 
		\item if $F$ is a field such that $K\subseteq F\subseteq E$ and 
			$f$ factorizes
			linearly in $F[X]$, then $F=E$. 
	\end{enumerate}
\end{definition}

Easy examples: 

\begin{example}
	$\C$ is a decomposition field of $X^2+1\in\R[X]$. 
\end{example}

\begin{example}
	$\Q[\sqrt{2}]$ is a decomposition field of $X^2-2\in\Q[X]$. 
\end{example}

\begin{example}
	$\Q(\sqrt[3]{2})$ is not a decomposition field of $X^3-2\in\Q[X]$. However, if
	$\omega$ ia a primitive cubic root of one, then 
	$\Q(\sqrt[3]{2},\omega)$ is is a decomposition field of $X^3-2\in\Q[X]$. 
\end{example}

\begin{proposition}
	$E$ is a decomposition field of $f\in K[X]$ if and only if
	$f$ factorizes linearly in $E[X]$ and $E=K(x_1,\dots,x_n)$ where 
	$x_1,\dots,x_n$ are roots of $f$. 
\end{proposition}

\begin{proof}
    Let $f=a\prod_{i=1}^r(X-x_i)^{n_i}$ and $F=K(x_1,\dots,x_r)$ with $x_1,\dots,x_r\in E$. Since $f$
    factorizes linearly in $F[X]$, it follows that $F=E$. 
    Conversely, let $E=K(x_1,\dots,x_r)$ and assume that $f$ factorizes linearly
    in $F[X]$. Then, in particular, $x_1,\dots,x_r\in F$. Hence $E\subseteq F$ and
    $F=E$. 
\end{proof}

One immediately obtains the following consequence:
If $E$ is a decomposition field of $f\in K[X]$, then $E/K$ is finite. 

\begin{theorem}
    Let $f\in K[X]$. There exists a (unique up to extension isomorphism) 
    decomposition field of $f$ over $K$. 
\end{theorem}

\begin{proof}
    Let $C/K$ be an algebraic closure. Write $f=a\prod_{i=1}^r(X-x_i)^{n_i}$ in $C[X]$. 
    Then $E=K(x_1,\dots,x_r)$ is a decomposition field of $f$ over $K$. Let us prove
    uniqueness: if $E_1/K$ is a decomposition field of $f$ over $K$, 
    then $E_1/K$ is algebraic and thus Proposition
    \ref{pro:Artin} implies that 
    there exists $\varphi\in\Hom(E_1/K,C/K)$ such that $\varphi|_K$ is the identity.
    Factorize $f$ linearly in $E_1[X]$ and apply $\overline{\varphi}$:
    \[
    f=a\prod_{j=1}^s(X-y_j)^{m_j}
    \implies
    \overline{\varphi}(f)=\varphi(a)\prod_{j=1}^s(X-\varphi(y_j))^{m_j}
    \]
    so $f$ factorizes linearly in $\varphi(E_1)$. Moreover, 
    $E_1=K(y_1,\dots,y_s)$ and it follows that 
    $\varphi(E_1)=K(\varphi(y_1),\dots,\varphi(y_s))$. Thus
    $\varphi(E_1)$ is a decomposition field of $f$. Since  
    $\varphi(E_1)\subseteq C$, it follows that $\varphi(E_1)=E$. 
\end{proof}

\begin{exercise}
If $E/K$ is finite and $\varphi\in\Hom(E/K,E/K)$, 
then $\varphi$ is an isomorphism. 
\end{exercise}

Let $C$ be an algebraic closure of $K$ and 
$G=\Gal(C/K)$. The group $G$ acts on $C$
\[
\sigma\cdot x=\sigma(x),\quad
\sigma\in G,\,x\in C.
\]
The orbits 
are of the form 
\[
O_G(x)=\{\sigma(x):\sigma\in G\}
=\{y\in C:y=\sigma(x)\text{ for some $\sigma\in G$}\}
\]
The elements $x,y\in C$ are \textbf{conjugate} 
if $y=\sigma(x)$ for some $\sigma\in G$. 

\begin{proposition}
    Let $C$ be an algebraic closure of $K$ and $x,y\in C$. Then 
    $x$ and $y$ are conjugate if and only if $f(x,K)=f(y,K)$. In particular, 
    $O_G(x)$ is finite. 
\end{proposition}

\begin{proof}
    Let $G=\Gal(C/K)$. 
    If $x$ and $y$ are conjugate, say $y=\sigma(x)$ for some $\sigma\in G$, 
    let us write $g=f(x,K)$ as 
    \[
    g=X^n+\sum_{i=0}^{n-1} a_iX^i. 
    \]
    Then $0=g(x)=x^n+\sum_{i=0}^{n-1}a_ix^i$ and hence $y$ is
    a root of $g$, as 
    \begin{align*}
    0=\sigma\left(x^n+\sum_{i=0}^{n-1}a_ix^i\right)
    &=\sigma(x)^n+\sum_{i=0}^{n-1}\sigma(a_i)\sigma(x)^i\\
    &=\sigma(x)^n+\sum_{i=0}^{n-1}a_i\sigma(x)^i
    =y^n+\sum_{i=0}^{n-1}a_iy^i.
    \end{align*}
    Thus $f(y,K)=g$. 
    
    Conversely, assume that $f(x,K)=f(y,K)$. Let
    $g=f(x,K)=f(y,K)$ and let 
    \[
    \varphi\colon K[x]\to K[y],
    \quad
    h(x)\mapsto h(y).
    \]
    Let us show that the map $\varphi$ is well-defined: we need to show 
    that if 
    $h_1(x)=h_2(x)$, then $h_1(y)=\varphi(h_1(x))=\varphi(h_2(x))=h_2(y)$. 
    If $h_1(x)=h_2(x)$, then 
    \[
    (h_1-h_2)(x)=h_1(x)-h_2(x)=0.
    \]
    Thus implies
    that $g$ divides $h_1-h_2$. In particular, $h_1(y)=h_2(y)$.
    
    A straightforward calculation shows that $\varphi$ is a field 
    homomorphism such that $\varphi|_K=\id$, so $\varphi$ is 
    an extension homomorphism such that $\varphi(x)=y$. There exists
    $\sigma\in\Hom(C/K,C/K)$ such that 
    $\sigma|_{K[x]}=\varphi$. Since $\sigma$ is a bijective, 
    $\sigma(x)=\varphi(x)=y$ and hence 
    $O_G(x)=O_G(y)$. 
\end{proof}

\begin{proposition}
    Let $C$ be an algebraic closure of $K$ and $x$. Then 
    \[
    f(x,K)=\prod_{y\in O_G(x)}(X-y)^m
    \]
    for some $m$. 
\end{proposition}

\begin{proof}
    For each $y\in O_G(x)$ let $m_y$ be the multiplicity
    of $y$ in $f(x,K)$. 
    Then, for example,
    $f(x,K)=(X-x)^{m_x}g$ for some $g$. If $y\in O_G(x)$, 
    then $y=\sigma(x)$ for some $\sigma\in\Gal(C/K)$. Since
    \[
    \overline{\sigma}(f(x,K))=f(x,K)=(X-y)^{m_x}\overline{\sigma}(g), 
    \]
    it follows that $m_y\geq m_x$. By symmetry, 
    we conclude that $m_x=m_y$. 
\end{proof}

The previous proposition shows, in particular, 
that all the roots of 
an irreducible polynomial $f\in K[X]$ 
in an algebraic closure $C$ of $K$
have the same multiplicity. This is clearly 
not true if $f$ is not irreducible. Find an example.

%Take 
%for example $f=(X-1)^2(X^2+1)\in\Q[X]$. 

% pag 20

\begin{definition}
\index{Decomposition field}
    Let $K$ be a field and $\{f_i:i\in I\}$ be a non-empty 
    family of polynomials of positive degree
    with coefficients in $K$. A \textbf{decomposition field} 
    of the $\{f_i:i\in I\}$ is an extension $E/K$
    such that every $f_i$ factorizes linearly in $E[X]$ and 
    if $F/K$ is a subextension of $E/K$ such that every $f_i$ 
    factorizes linearly in $F[X]$, then $F=E$. 
\end{definition}

\begin{exercise}
    Prove that $E/K$ is a decomposition field of
    $\{f_i:i\in I\}$ if and only if every $f_i$ factorizes linearly 
    in $E[X]$ and $E=K(S)$ where $S=\{\text{roots of $f_i$ for all $i$}\}$. 
\end{exercise}

\begin{exercise}
    Prove that if $E/K$ is a decomposition field
    of $\{f_i:i\in I\}$, then $E/K$ is algebraic. If, moreover, 
    $I$ is finite, then $E/K$ is a decomposition field
    of $\prod_{i\in I}f_i$. 
\end{exercise}

\begin{exercise}
    Prove that 
    there exists a decomposition field of $\{f_i:i\in I\}$ 
    and it is unique up to extension isomorphism. 
\end{exercise}

\topic{Normal extensions}

\begin{proposition}
    Let $E/K$ be an algebraic extension and $\sigma\in\Hom(E/K,E/K)$. 
    Then $\sigma$ is bijective. 
\end{proposition}

\begin{proof}
    Let $x\in E$ and 
    $C$ be an algebraic closure of $K$ that contains $E$. There
    exists $\varphi\colon C\to C$ such that $\varphi|_E=\sigma$. 
    Thus $\varphi|_K=\sigma|_K=\id_K$. Let $G=\Gal(C/K)$. Then  
    $\varphi\in G$. If $z\in O_G(x)$, 
    then $z=\tau(x)$ for some $\tau\in G$ and hence
    \[
    \varphi(z)=\varphi(\tau(x))=(\varphi\tau)(x). 
    \]
    This implies that $\varphi(z)\in O_G(x)$ 
    and $\varphi(O_G(x))=O_G(x)$. Thus $\sigma|_{(E\cap O_G(x)}$ is injective, as 
    \begin{align*}
    \sigma(E\cap O_G(x))&=\varphi(E\cap O_G(x))\\
    &=\varphi(E)\cap\varphi(O_G(x))
    =\sigma(E)\cap O_G(x)\subseteq E\cap O_G(x).
    \end{align*}
    Since $|E\cap O_G(x)|<\infty$, it follows that $E\cap O_G(x)=\sigma(E\cap O_G(x))$ and
    hence $x$ belongs to the image of $\sigma$. 
\end{proof}

% pag 21 example

\begin{definition}
    Let $E/K$ be an algebraic extensions and $C$ be an algebraic closure of $K$. 
    Then $E/K$ is \textbf{normal} if 
    $\sigma(E)\subseteq E$ for all $\sigma\in\Hom(E/K,C/K)$. 
\end{definition}

Note that $\sigma(E)\subseteq E$ in the previous definition
is equivalent to $\sigma(E)=E$. 

\begin{example}
    The extension $\Q(\sqrt[3]{2})/\Q$ is not normal. Why?
\end{example}

Some trivial examples of normal extensions: $K/K$ is normal
and if $C$ is an algebraic closure of $K$, then $C/K$ is normal. 

\begin{example}
    The extension $\Q(\sqrt{2})/\Q$ is normal. In fact, 
    every extension generated by algebraic elements of degree two is normal. 
\end{example}

\begin{exercise}
    Let $\xi$ be a primitive cubic root of one. Then 
    $\Q(\sqrt[3]{2},\xi)/\Q$ is normal. 
\end{exercise}

The following result is useful but technical, that is why we leave the proof
as an exercise. 

\begin{exercise}
    Prove that the previous definition depends on $E$ and not on the
    algebraic closure $C$. 
\end{exercise}

Some properties:

\begin{proposition}
    Let $E/K$ be a normal extension and $f\in K[X]$ be an irreducible polynomial
    that admits a root $x$ in $E$. Then $f$ factorizes
    linearly in $E$.
\end{proposition}

\begin{proof}
    We may assume that $f$ is monic. Let $C/K$ be an algebraic closure of $K$ containing $E$. 
    Let $y$ be a root of $f$ in $C$. Since $f=f(x,K)=f(y,K)$, 
    it follows that $y=\sigma(x)$ for some $\sigma\in\Gal(C/K)$. Since 
    $E/K$ is normal, $\sigma|_E\colon E\to C$ is an automorphism of $E/K$, that is
    $\sigma(E)\subseteq E$. In particular, $y\in E$. 
\end{proof}

Let $K\subseteq F\subseteq E$ be a tower of fields. 
If $E/K$ is normal, then $E/F$ is normal. However, 
Note that $E/K$ normal does not imply $F/K$ normal, as this would imply 
that every extension is normal. Moreover, 
$E/K$ normal and $F/K$ normal do not imply $E/K$ normal.
    
\begin{example}
The extensions $\Q(\sqrt[4]{2})/\Q(\sqrt{2})$ and $\Q(\sqrt{2})/\Q$ are both
normal, but $\Q(\sqrt[4]{2})/\Q$ is not normal, 
as the roots of $X^4-2$ are
$\sqrt{2}$, $-\sqrt{2}$, $\sqrt{2}i$ and $-\sqrt{2}i$.
\end{example}


Recall that if $C$ is an algebraic closure of $K$ and $x\in C$,
then 
\[
f(x,K)=\prod(X-y)^m,
\]
where the product is taken over all 
$y\in O_{\Gal(C/K)}(x)$. 
If $E/K$ is normal and $x\in E$, then there exists $m$ such that 
\[
f(x,K)=\prod(X-y)^m,
\]
where the product is taken over all 
$y\in O_{\Gal(E/K)}(x)$. 

\begin{proposition}
    Let $K\subseteq F\subseteq E$ be a tower of fields. If $F/K$ 
    is normal, then $EF/E$ is normal.
\end{proposition}

\begin{proof}
    Let $C$ be an algebraic closure of $E$ containing $EF$. 
    Let $\sigma\in\Hom(EF/E,C/E)$. We claim that $\sigma(EF)=EF$. Let 
    \[
    \overline{K}=\{x\in C:x\text{ is algebraic over $K$}\}.
    \]
    Then $\overline{K}$ is an algebraic closure over $K$ and $F\subseteq\overline{K}$. 
    Since $F/K$ is normal and $\sigma|_F\in\Hom(F/K,\overline{K}/K)$, 
    it follows that $\sigma(F)=F$. If $z\in EF$, then
    $z=\sum_{i=1}^m e_if_i$ for some $e_1,\dots,e_m\in E$ and 
    $f_1,\dots,f_m\in F$. Then 
    \[
    \sigma(z)=\sum_{i=1}^m\sigma(e_i)\sigma(f_i)=\sum_{i=1}^m e_i\sigma(f_i)\in EF.\qedhere 
    \]
\end{proof}

\begin{proposition}
    Let $E/K$ be an algebraic extension. Then 
    $E/K$ is normal if and only if $E/K$ is the decomposition field
    of a family of polynomials of $K[X]$ of positive degree.
\end{proposition}

\begin{proof}
    Let $G=\Gal(E/K)$. If $x\in E$ and $f(x,K)=\prod_{y\in O_G(x)}(X-y)^m$, 
    then $f(x,K)$ factorizes linearly in $E[X]$. Thus 
    $E/K$ is a decomposition field of the family 
    $\{f(x,K):x\in E\}$. 
    Conversely, assume that $E/K$ is a decomposition field of the family 
    $\{f_i:i\in I\}$. Then $E=K(S)$ where $S$ is the set of roots
    of the polynomials $f_i$. Let $C/K$ be an algebraic closure
    of $K$ that contains $E$ and let $\sigma\in\Hom(C/K,E/K)$. Let $x\in S$. 
    Then $x$ is a root of some $f_j=\sum a_kX^k$. Since $f_j(x)=0$, 
    it follows that $\sigma(x)$ is a root of $f_j$, as 
    \[
    f_j(\sigma(x))=\sum a_k\sigma(x)^k
    =\sum\sigma(a_k)\sigma(x^k)
    =\sigma\left(\sum a_kx^k\right)=\sigma(0)=0.
    \]
    Hence $\sigma(E)\subseteq E$. 
\end{proof}

\section{Dedekind's theorem}

Note that every extension homomorphism $E/K\to F/K$ is, in particular, 
a $K$-linear map $E\to F$, that is
\[
\Hom(E/K,F/K)\subseteq\Hom_K(E,F).
\]
If $F/K$ is an extension and $V$ 
is a $K$-vector space, the set
$\Hom_K(E,F)$ of $K$-linear maps
is a vector space over $F$ with
$(a\cdot f)(v)=af(v)$ for $a\in F$, $f\in\Hom_K(E,F)$ and $v\in V$. 

\begin{exercise}
\label{xca:dim}
    Prove that $\dim_F\Hom_K(V,F)\geq\dim_KV$. Moreover, if 
     $\dim_KV<\infty$, then $\dim_F\Hom_K(V,F)=\dim_KV$.
\end{exercise}

\begin{theorem}[Dedekind]
\index{Dedekind's theorem}
Let $E/K$ and $F/K$ be extensions and let 
$\{\varphi_i:i\in I\}$ 
be a subset of
$\Hom(E/K,F/K)$, i.e. a 
family of extension homomorphisms. Assume that 
$\varphi_i\ne \varphi_j$ if $i\ne j$. Then 
the subset $\{\varphi_i:i\in I\}\subseteq\Hom_K(E,F)$ 
is linearly independent over $F$. 
\end{theorem}

\begin{proof}
    Assume it is not. Let $\{\varphi_1,\dots,\varphi_n\}$ 
    be linearly dependent over $F$ with $n$ minimal. Clearly, $n>1$. 
\end{proof}

    